---
title: "STM_Phase1_Final"
author: "Moira"
date: "2026-02-04"
output: html_document
---

In this notebook, I run the Structural Topic Modelling (STM) Analysis along with the Krippendorf check, and execute the Multivariate Change-Point Analysis and the multivariate Mixed Effects Model. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# clear
rm(list = ls())
```

# Load Data & Clean

```{r}
library(here)
library(stm)
library(tm)
library(readr)
library(dplyr)
library(gt)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(igraph)
library(lubridate)
library(irr)
library(zoo)
library(ecp)
library(lme4)
library(lmerTest)  # gives p-values for lmer
library(broom.mixed)

dir.create("R_Figures")
dir.create("R_Tables")

df_base <- read_csv("data/parliamentary_baseline_cleaned.csv")
df_MandB <- read_csv("data/df_merged_clean.csv")

REFIT_MODELS <- FALSE
REFIT_BASELINE <- FALSE
```
baseline_df = pd.read_csv("/Users/moira/Documents/CEU/Thesis/1_Data_Collection/parliamentary_baseline_cleaned.csv")
df_merged_clean = pd.read_csv('/Users/moira/Documents/CEU/Thesis/2_Data_Cleaning/df_merged_clean.csv')


```{r}
df_MandB$party[is.na(df_MandB$party)] <- "Guest Speaker"
df_MandB$in_government[df_MandB$party == "Guest Speaker"] <- 0
```


```{r}
# Irish language stopwords for Dáil corpus
irish_stopwords <- c(
  "agus", "seo", "atá", "bhí", "bhfuil", "bheith",
  "chur", "daoin", "ansin", "agam", "siad", "acu",
  "níl", "níos", "déanta", "rá", "teachtaí",
  "nach", "más", "faoi", "mbeadh", "anoi",
  "tánaiste", "oireachtas"
) #claude helped me come up with these

# helper function to clean for stm
prepare_for_stm <- function(df) {
  df$date <- as.Date(df$date)
  df$date_numeric <- as.numeric(df$date)
  df$party <- as.factor(df$party)
  df$in_government <- as.factor(df$in_government)
  df <- df[!is.na(df$party) & !is.na(df$in_government), ]
  
  processed <- textProcessor( #already includes english stopwords in the apckage
    df$clean_text, 
    metadata = df,
    customstopwords = irish_stopwords # removes Irish language tokens
  ) # comes from STM package
  
  out <- prepDocuments(
    processed$documents, 
    processed$vocab, 
    processed$meta, 
    lower.thresh = 5 # removes words in fewer than 5 documents, converts words into matrix format, returns a list for stm()
  ) 
  
  return(out)
}

out_mbh  <- prepare_for_stm(df_MandB) # lists with stm elements
out_base <- prepare_for_stm(df_base)
```
^ stopword removal and frequency filtering of 5 can result in 24 documents being removed due to "No Words"




# Phase 1
##---------- MBH ---------- 

## MBH STM

```{r}
if (REFIT_MODELS) {
  set.seed(42)
  custom_k <- c(5, 7, 9, 11, 13, 15, 17, 20, 22, 25, 30, 40, 50)
  diag_search <- searchK(out_mbh$documents, out_mbh$vocab, 
                         K = custom_k,
                         prevalence =~ party + in_government + s(date_numeric),
                         data = out_mbh$meta, 
                         cores = 1)
  saveRDS(diag_search, "R_data/mbh_diagnostic_results.rds")
} else {
  diag_search <- readRDS("R_data/mbh_diagnostic_results.rds")
}
print(diag_search$results)

```



```{r}
# plot of results
plot(diag_search)

```


```{r}
png("R_Figures/diag_search_plot_mbh.png", width = 800, height = 600)

# create the plot
plot(diag_search)

# close device
dev.off()

```



The final model with the best k = 22

```{r}
if (REFIT_MODELS) {
  set.seed(42)
  mbh_select <- selectModel(out_mbh$documents, out_mbh$vocab,
                            K = 22,
                            prevalence =~ party + in_government + s(date_numeric),
                            data = out_mbh$meta,
                            runs = 20,
                            max.em.its = 150, #default 75 insufficient; model converged at ~94 iterations
                            init.type = "Spectral")
  saveRDS(mbh_select, "R_data/mbh_select_22.rds")
} else {
  mbh_select <- readRDS("R_data/mbh_select_22.rds")
}

# Save to Figures folder
png("R_Figures/plotModels_mbh_22.png", width = 800, height = 600)
plotModels(mbh_select, pch = 1:20, legend.position = "bottomright")
dev.off()

# Display in RStudio plots pane
plotModels(mbh_select, pch = 1:20, legend.position = "bottomright")

# !! STOP HERE: look at plot, note best run number (top-right, tightest spread)
```
```{r}
# Get exclusivity and coherence scores for each surviving run
lapply(mbh_select$runout, function(m) {
  list(
    exclusivity = mean(exclusivity(m)),
    coherence = mean(semanticCoherence(m, out_mbh$documents))
  )
})

```

All runs are identical and converged to the same solution every time. Robust! Slightly redundant with spectral initialization but a good robustness check.

```{r}
# After inspecting plotModels, setting the best run number here
mbh_stm <- mbh_select$runout[[1]]  # <-- doesn't matter as all were the same

saveRDS(mbh_stm, "R_data/mbh_stm_22.rds")
saveRDS(out_mbh, "R_data/out_mbh_22.rds")

```

```{r}
labelTopics(mbh_stm)

```


```{r}
topic_labels <- c(
  "T1"  = "T1: Mother and Baby Homes & County Homes",
  "T2"  = "T2: Survivor Redress & Exclusions",
  "T3"  = "T3: Party Politics",
  "T4"  = "T4: Tuam Site & Excavation",
  "T5"  = "T5: Childcare Funding",
  "T6"  = "T6: Human Rights & Legislation",
  "T7"  = "T7: Commission Reports & Delays",
  "T8"  = "T8: Redress Administration",
  "T9"  = "T9: Adoption & Identity",
  "T10" = "T10: Archives & Records",
  "T11" = "T11: Infant Mortality & Vaccine Trials",
  "T12" = "T12: Family Law & Social Norms",
  "T13" = "T13: Parliamentary Procedure",
  "T14" = "T14: Order of Business & Seanad",
  "T15" = "T15: Abortion & Eighth Amendment",
  "T16" = "T16: Informal & Direct Debate",
  "T17" = "T17: Commission Scope & Establishment",
  "T18" = "T18: Grace Case & Child Protection",
  "T19" = "T19: Survivor Truth & Shame",
  "T20" = "T20: Formal & Supportive Debate",
  "T21" = "T21: Irish Language",
  "T22" = "T22: Church Role"
)

```

```{r}
# Find example speeches
findThoughts(mbh_stm, 
             texts = out_mbh$meta$clean_text,
             topics = c(1:22),
             n = 3)  # 3 example speeches per topic

```



## Descriptive Tables

Make some descriptive tables that show how topics were talked about over time.... and which parties.... etc, topic prevalence changes over time... 


PARTY DOMINANCE TABLE
```{r}
# create the Party Summary Table
party_dominance <- out_mbh$meta %>%
  group_by(party) %>%
  summarise(
    Speech_Count = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    Percentage = round((Speech_Count / sum(Speech_Count)) * 100, 2)
  ) %>%
  arrange(desc(Speech_Count))

# view the table
print(party_dominance)

# Save xlsx
write_xlsx(party_dominance, "R_Tables/party_dominance_table.xlsx")

# save R version
saveRDS(party_dominance, "R_Tables/party_dominance_table.rds")


```

PARTY DOMINANCE PLOT
```{r}
# plotting the dominance of party
ggplot(party_dominance, aes(x = reorder(party, Speech_Count), y = Speech_Count, fill = Speech_Count)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.7) +
  coord_flip() +
  theme_minimal(base_size = 12) +
  labs(title = "Corpus Dominance by Party (MBH Dataset)",
       subtitle = "Total speech count per political grouping",
       x = "Political Party", 
       y = "Total Number of Speeches") +
  theme(legend.position = "none")

ggsave("R_Figures/party_dominance_plot.png", width = 8, height = 6, dpi = 300)


```

COALITION MEMBERS WHEN TIMELINE
```{r}
# identify distinct government status for each party
timeline_chapters <- df_MandB %>%
  mutate(date = as.Date(date)) %>%
  arrange(party, date) %>%
  group_by(party) %>%
  # data.table::rleid creates a unique ID every time 'in_government' changes
  mutate(streak_id = data.table::rleid(in_government)) %>% 
  group_by(party, streak_id, in_government) %>%
  summarise(
    Start_Date = min(date),
    End_Date = max(date),
    Speech_Count = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    Status = ifelse(in_government == 1, "Government", "Opposition")
  ) %>%
  arrange(Start_Date)

# create the table
pretty_chapters_table <- timeline_chapters %>%
  select(party, Status, Start_Date, End_Date, Speech_Count) %>%
  gt(groupname_col = "Status") %>% 
  tab_header(
    title = md("**Institutional Chapters of the MBH Corpus**"),
    subtitle = "Timeline showing distinct role transitions per party"
  ) %>%
  # FIX: Using "yMMM" for Month Year (e.g., Nov 2010)
  fmt_date(columns = c(Start_Date, End_Date), date_style = "yMMM") %>%
  cols_label(
    party = "Political Party",
    Start_Date = "Joined",
    End_Date = "Exited",
    Speech_Count = "Speeches"
  ) %>%
  tab_options(
    row_group.background.color = "#f9f9f9",
    row_group.font.weight = "bold"
  )

# View the table
pretty_chapters_table

# Save as PNG
pretty_chapters_table %>%
  gtsave(filename = "R_Tables/Institutional_Chapters_Table.png")

# Save as PDF
pretty_chapters_table %>%
  gtsave(filename = "R_Tables/Institutional_Chapters_Table.pdf")

print("Table saved to the Tables folder!")
```




BALANCE CHECK
```{r}
# Create the 3 groups
status_breakdown <- df_MandB %>%
  mutate(
    # Force column to character to check for the label
    party_char = as.character(party),
    # If party is Guest Speaker, label them as such. Otherwise, split by government status.
    Final_Category = case_when(
      party_char == "Guest Speaker" ~ "Guest Speaker",
      in_government == 1 ~ "Government",
      in_government == 0 ~ "Opposition",
      TRUE ~ "Other"
    )
  ) %>%
  group_by(Final_Category) %>%
  summarise(Speech_Count = n(), .groups = 'drop') %>%
  mutate(
    Percentage = (Speech_Count / sum(Speech_Count)) * 100
  ) %>%
  arrange(desc(Speech_Count))

# Create the GT table
corpus_composition_table <- status_breakdown %>%
  gt() %>%
  tab_header(
    title = md("**Corpus Composition by Executive Status**"),
    subtitle = "Includes Guest Speakers as a distinct non-parliamentary category"
  ) %>%
  fmt_number(
    columns = Percentage,
    decimals = 1,
    pattern = "{x}%"
  ) %>%
  cols_label(
    Final_Category = "Speaker Role",
    Speech_Count = "Total Speeches",
    Percentage = "Proportion"
  ) %>%
  tab_options(
    table.width = px(450),
    column_labels.font.weight = "bold"
  )

# Save
corpus_composition_table %>%
  gtsave(filename = "R_Tables/Detailed_Corpus_Composition.png")

# View result
print(status_breakdown)
corpus_composition_table

```

TOPIC PREVALENCE PLOT
```{r}
# extract and average the theta values
agg_prevalence <- data.frame(colMeans(mbh_stm$theta))
colnames(agg_prevalence) <- "Mean_Prevalence"
agg_prevalence$Topic_ID <- paste0("T", 1:22)
agg_prevalence$Full_Label <- topic_labels[agg_prevalence$Topic_ID]

# create a plot
ggplot(agg_prevalence, aes(x = reorder(Full_Label, Mean_Prevalence), 
                           y = Mean_Prevalence)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.7) +
  coord_flip() +
  theme_minimal(base_size = 12) +
  labs(title = "Aggregate Topic Prevalence (MBH Dataset)",
       x = "Topic", 
       y = "Average Proportion in Corpus")

ggsave("R_Figures/agg_topic_prevalence.png", width = 8, height = 6, dpi = 300)


```


TOPIC PREVALENCE PROFILES BY POLITICAL PARTY
Excluded: Socialist Party & WUA
```{r}
# specify parties to exclude
exclude_list <- c("socialist party", "workers and unemployed action")

# prepare party-level data
party_prevalence <- data.frame(mbh_stm$theta)
colnames(party_prevalence) <- paste0("T", 1:22)
party_prevalence$party <- out_mbh$meta$party

# clean and filter data
party_long <- party_prevalence %>%
  mutate(party_clean = tolower(trimws(as.character(party)))) %>%
  filter(!party_clean %in% exclude_list, party_clean != "") %>%
  mutate(party = droplevels(as.factor(party_clean))) %>%
  group_by(party) %>%
  summarise(across(starts_with("T"), mean)) %>%
  pivot_longer(cols = starts_with("T"), names_to = "Topic_ID", values_to = "Prevalence") %>%
  mutate(
    Full_Label = topic_labels[Topic_ID],
    Full_Label_Wrapped = str_wrap(Full_Label, width = 25),
    Topic_Num = as.numeric(str_extract(Topic_ID, "[0-9]+"))
  )

# create facet wrap by party
p <- ggplot(party_long, aes(x = reorder(Full_Label_Wrapped, Topic_Num), 
                             y = Prevalence, fill = Prevalence)) +
  geom_bar(stat = "identity") +
  facet_wrap(~party, scales = "free_x", ncol = 3) +  # fixed x, scroll topics on y
  coord_flip() +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  theme_minimal(base_size = 9) +
  labs(title = "Topic Prevalence Profiles by Political Party",
       subtitle = "Excluded: Socialist Party & WUA",
       x = NULL, y = "Average Prevalence") +
  theme(
    legend.position = "none",
    axis.text.y = element_text(size = 6),
    axis.text.x = element_text(size = 6),
    strip.text = element_text(face = "bold", size = 8),
    panel.spacing = unit(0.8, "lines")
  )

ggsave("R_Figures/party_discursive_profile.png", plot = p, width = 16, height = 20, dpi = 300)
ggsave("R_Figures/party_discursive_profile.pdf", plot = p, width = 16, height = 20)

```

Topic proportions by PArty Table
```{r}
# extract Topic Proportions and merge with Party
topic_party_dist <- data.frame(mbh_stm$theta)
colnames(topic_party_dist) <- paste0("Topic", 1:22)
topic_party_dist$party <- out_mbh$meta$party

# calculate mean prevalence per party
party_table <- topic_party_dist %>%
  group_by(party) %>%
  summarise(across(starts_with("Topic"), mean))

# transpose for easier reading
party_table_t <- as.data.frame(t(party_table[,-1]))
colnames(party_table_t) <- party_table$party
party_table_t$Topic <- rownames(party_table_t)

print(party_table_t)

```


## MBH Estimate Effect - Topic Ownership Plot

TOPIC OWNERSHIP (GOV vs OPP)
```{r}
# estimate the effect of Government status on all 15 topics
prep_gov <- estimateEffect(1:22 ~ in_government, mbh_stm, meta = out_mbh$meta)



# plot the difference
#pdf("R_Figures/topic_ownership_difference.pdf", width = 12, height = 8)
png("R_Figures/topic_ownership_gov_vs_opp.png",
    width = 3000, height = 2000, res = 300)


par(mar = c(5, 12, 4, 4),
    cex = 0.6)         # all other text including topic labels

plot(prep_gov, "in_government", method = "difference",
     cov.value1 = "1", cov.value2 = "0",
     main = "Topic Prevalence: Opposition vs. Government",
     xlab = "Opposition <---> Government",
     labeltype = "custom",
     custom.labels = topic_labels
)


```

```{r}
summary(prep_gov)
```

```{r}
names(mbh_stm)
```


CHECK IF CONVERGED
```{r}
mbh_stm$convergence
```


MBH TOPIC OWNERSHIP SHARE DECOMPOSIITION (%)
```{r}
## Regime Share Decomposition: Topic Ownership by % (Gov vs Opp)
regime_share_data <- data.frame(
  mbh_stm$theta,
  party        = out_mbh$meta$party,
  in_government = as.numeric(as.character(out_mbh$meta$in_government)),
  date         = as.Date(out_mbh$meta$date)
) 
colnames(regime_share_data)[1:22] <- paste0("T", 1:22)

# Add time period labels
regime_share_data <- regime_share_data %>%
  mutate(
    period = case_when(
      date < as.Date("2021-01-12") ~ "Pre-Report (before Jan 2021)",
      TRUE                         ~ "Post-Report (Jan 2021+)"
    ),
    role = ifelse(in_government == 1, "Government", "Opposition")
  )

# Function: compute % of each topic's total usage coming from Gov vs Opp
compute_share <- function(df, label) {
  df %>%
    pivot_longer(cols = T1:T22, names_to = "Topic_ID", values_to = "Proportion") %>%
    group_by(Topic_ID, role) %>%
    summarise(Total = sum(Proportion), .groups = "drop") %>%
    group_by(Topic_ID) %>%
    mutate(
      Grand_Total = sum(Total),
      Share_Pct   = (Total / Grand_Total) * 100,
      Period      = label
    ) %>%
    ungroup()
}

# Run for all 3 windows
share_all    <- compute_share(regime_share_data,                                    "All Time")
share_pre21  <- compute_share(filter(regime_share_data, period == "Pre-Report (before Jan 2021)"),  "Pre-2021")
share_post21 <- compute_share(filter(regime_share_data, period == "Post-Report (Jan 2021+)"), "Post-2021")

share_combined <- bind_rows(share_all, share_pre21, share_post21) %>%
  mutate(
    Full_Label = topic_labels[Topic_ID],
    # Flag regime membership for ordering
    Regime = case_when(
      Topic_ID %in% c("T4","T5","T8","T13","T14","T17", "T20") ~ "Managerial",
      Topic_ID %in% c("T2","T3","T6","T11","T12","T16","T18", "T19","T21", "T22") ~ "Exploratory",
      TRUE ~ "Other"
    )
  )

# Stacked bar: Gov share vs Opp share per topic (faceted by period)
ggplot(share_combined, aes(x = reorder(Full_Label, Share_Pct * (role == "Government")),
                           y = Share_Pct, fill = role)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_hline(yintercept = 50, linetype = "dashed", color = "black", linewidth = 0.5) +
  coord_flip() +
  facet_wrap(~factor(Period, levels = c("All Time","Pre-2021","Post-2021"))) +
  scale_fill_manual(values = c("Government" = "#2166AC", "Opposition" = "#D6604D")) +
  theme_minimal() +
  labs(
    title    = "Topic Prevalence: Government vs. Opposition Share (%)",
    x = NULL, y = "Share of Total Topic Usage (%)",
    fill = "Role",
    caption = "Managerial: T4,T5,T8,T13,T14,T17,T20 | Exploratory: T2,T3,T6,T11,T12,T16,T18,T19,T21,T22"
  ) +
  theme(
    axis.text.y  = element_text(size = 7),
    strip.text   = element_text(face = "bold"),
    legend.position = "bottom"
  )

ggsave("R_Figures/topic_ownership_share_pct.png", width = 14, height = 8, dpi = 300)
ggsave("R_Figures/topic_ownership_share_pct.pdf", width = 14, height = 8)

```
```{r}
## Clean numeric summary table of the shifts

share_wide <- share_combined %>%
  filter(role == "Government") %>%
  select(Topic_ID, Full_Label, Regime, Period, Share_Pct) %>%
  pivot_wider(names_from = Period, values_from = Share_Pct) %>%
  mutate(
    Shift = `Post-2021` - `Pre-2021`,
    across(where(is.numeric), ~round(.x, 1))
  ) %>%
  arrange(desc(`All Time`))

print(share_wide)
write_xlsx(share_wide, "R_Tables/topic_ownership_share_table.xlsx")
```



REGIME SCHEMA PLOT
```{r}
regime_schema <- data.frame(
  Regime = c(
    rep("Managerial", 7),
    rep("Exploratory", 10)
  ),
  Topic = c(
    # Managerial - government owned from ownership plot
    "T4: Tuam Site & Excavation",
    "T5: Childcare Funding",
    "T8: Redress Administration",
    "T13: Parliamentary Procedure",
    "T14: Order of Business & Seanad",
    "T17: Commission Scope & Establishment",
    "T20: Formal & Supportive Debate",
    # Exploratory - opposition owned from ownership plot
    "T2: Survivor Redress & Exclusions",
    "T3: Party Politics",
    "T6: Human Rights & Legislation",
    "T11: Infant Mortality & Vaccine Trials",
    "T12: Family Law & Social Norms",
    "T16: Informal & Direct Debate",
    "T18: Grace Case & Child Protection",
    "T19: Church, State & Survivor Truth",
    "T21: Irish Language",
    "T22: Church & Hospital Ownership"
  ),
  FREX_Words = c(
    # Managerial FREX from new labelTopics output
    "excav, site, coron, burial, director, dna, forens",
    "childcar, alloc, fund, million, increas, earli, programm",
    "payment, card, applic, resid, waiver, elig, quirk",
    "shall, minut, busi, stage, exceed, seanad, statement",
    "transport, byrn, notic, senat, mcgrath, burk, pat",
    "establish, refer, scope, term, inquiri, examin, judg",
    "welcom, thank, import, stori, everyon, colleagu, listen",
    # Exploratory FREX from new labelTopics output
    "redress, exclud, survivor, trauma, six, arbitrari, exclus",
    "fine, gael, fianna, fáil, polit, taoiseach, parti",
    "human, right, court, principl, bill, law, jurisdict",
    "bethani, trial, mortal, vaccin, infant, die, unmark",
    "daughter, love, father, marri, lone, marriag, famili",
    "get, point, say, situat, thing, know, tri",
    "grace, asylum, homeless, mental, complaint, sexual, defenc",
    "victim, shame, truth, justic, horror, incarcer, apolog",
    "lei, sin, mar, narrat, ach, utter, read",
    "hospit, matern, chariti, etho, church, religi, ownership"
  )
)

regime_table <- regime_schema %>%
  gt(groupname_col = "Regime") %>%
  tab_header(
    title = "Table 2: The Regime Schema",
    subtitle = "Categorization of STM Topics by Linguistic and Rhetorical Function"
  ) %>%
  cols_label(
    Topic = "Topic ID & Label",
    FREX_Words = "Top FREX Words"
  ) %>%
  cols_width(
    Topic ~ px(220),
    FREX_Words ~ px(320)
  ) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 16,
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    row_group.background.color = "#f9f9f9"
  )

regime_table

# Save
regime_table %>%
  gtsave(filename = "R_Tables/Table2_Regime_Schema.png", expand = 10)
regime_table %>%
  gtsave(filename = "R_Tables/Table2_Regime_Schema.pdf")

```


## Co-occurence of Topics
CHECK WHICH TOPIC CO-OCCUR
PLOT
```{r}
topic_cor <- topicCorr(mbh_stm)
topic_cor$cor
plot(topic_cor)
plot(topic_cor, type = "network", cutoff = 0.1)
```
```{r}
png("R_Figures/topic_correlation_network.png", width = 3000, height = 3000, res = 300)

plot(topic_cor, type = "network", cutoff = 0.1)

dev.off()

```

## --------- Baseline --------- 

## Baseline STM

```{r}
if (REFIT_BASELINE) {
  set.seed(42)
  custom_k <- c(5, 10, 15, 20, 25, 30)
  diag_search_base <- searchK(out_base$documents, out_base$vocab, K = custom_k,
                         prevalence =~ party + in_government + s(date_numeric),
                         data = out_base$meta,
                         verbose = TRUE,
                         cores = 1)
  saveRDS(diag_search_base, "R_data/base_diagnostic_results.rds")
} else {
  diag_search_base <- readRDS("R_data/base_diagnostic_results.rds")
}
print(diag_search_base$results)
plot(diag_search_base)



```



```{r}
if (REFIT_BASELINE) {
  set.seed(42)
  base_select <- selectModel(out_base$documents, out_base$vocab,
                             K = 20,
                             prevalence =~ party + in_government + s(date_numeric),
                             data = out_base$meta,
                             runs = 20,
                             max.em.its = 150, #default 75 insufficient; model converged at ~94 iterations
                             init.type = "Spectral")
  saveRDS(base_select, "R_data/base_select.rds")
} else {
  base_select <- readRDS("R_data/base_select.rds")
}

# Save to Figures folder
png("R_Figures/plotModels_base.png", width = 800, height = 600)
plotModels(base_select, pch = 1:20, legend.position = "bottomright")
dev.off()

# Display
plotModels(base_select, pch = 1:20, legend.position = "bottomright")

# !! STOP HERE: look at plot, note best run number (top-right, tightest spread)

```


```{r}
# Get exclusivity and coherence scores for each surviving run
lapply(base_select$runout, function(m) {
  list(
    exclusivity = mean(exclusivity(m)),
    coherence = mean(semanticCoherence(m, out_base$documents))  )
})

```

```{r}
# After inspecting plotModels, setting the best run number here
  base_stm <- base_select$runout[[1]]  # <-- same results
  
  saveRDS(base_stm, "R_data/base_stm.rds")
  saveRDS(out_base, "R_data/out_base.rds")

```


```{r}
labelTopics(base_stm)
```

Baseline Labels
```{r}
base_labels <- c(
  "T1"  = "T1: Legislative Amendments",
  "T2"  = "T2: Gender & Human Rights",
  "T3"  = "T3: Bill Debates",
  "T4"  = "T4: Infrastructure & Defence",
  "T5"  = "T5: Local Government & Irish Language",
  "T6"  = "T6: Health & Hospital Services",
  "T7"  = "T7: Banking & Pensions",
  "T8"  = "T8: Family & Child Welfare",
  "T9"  = "T9: Parliamentary Questions",
  "T10" = "T10: Local Issues & Roads",
  "T11" = "T11: Mortgage & Farming Debt",
  "T12" = "T12: Banking & Finance",
  "T13" = "T13: Housing & Rent",
  "T14" = "T14: Public Administration & Garda",
  "T15" = "T15: General Political Debate",
  "T16" = "T16: Foreign Affairs & Brexit",
  "T17" = "T17: Arts & Prison Services",
  "T18" = "T18: Social Welfare & Payments",
  "T19" = "T19: Seanad Procedure",
  "T20" = "T20: Education & Schools"
)

```


## Descriptive Tables

Topic Prevalence Profiles by Political Party Baseline
```{r}

# process Baseline Data
base_prevalence <- data.frame(base_stm$theta)
colnames(base_prevalence) <- paste0("T", 1:20)
base_prevalence$party <- out_base$meta$party

base_long <- base_prevalence %>%
  filter(!party %in% exclude_list) %>%
  mutate(party = droplevels(as.factor(party))) %>%
  group_by(party) %>%
  summarise(across(starts_with("T"), mean)) %>%
  pivot_longer(-party, names_to = "Topic_ID", values_to = "Prevalence")

# plot Baseline Profiles
ggplot(base_long, aes(x = reorder(Topic_ID, Prevalence), y = Prevalence, fill = Topic_ID)) +
  geom_bar(stat = "identity") +
  facet_wrap(~party, scales = "free_y") + 
  coord_flip() +
  scale_x_discrete(labels = base_labels) + 
  theme_minimal() +
  labs(title = "Topic Prevalence Profiles by Political Party Baseline", subtitle = "Standard Political Business") +
  theme(legend.position = "none", axis.text.y = element_text(size = 7))


```

## Baseline Estimate Effect - Topic Ownership Plot
Topic Ownership
```{r}
# estimate the effect of all topics
prep_base_gov <- estimateEffect(1:20 ~ in_government, base_stm, meta = out_base$meta)

par(mar = c(5, 12, 4, 2)) 

plot(prep_base_gov, "in_government", method = "difference",
     cov.value1 = "1", cov.value2 = "0",
     topics = 1:20,
     main = "Baseline: Topic Ownership (Gov vs. Opp)",
     xlab = "Opposition <--- (Difference) ---> Government",
     labeltype = "custom",
     custom.labels = base_labels)

```

## Co-occurence of Topics
```{r}
topic_cor <- topicCorr(base_stm)
topic_cor$cor
plot(topic_cor)
plot(topic_cor, type = "network", cutoff = 0.1)
```

# Phase 2

## Table of Mean Topic Prevalence by Year
```{r}
# create a Year variable in metadata
out_mbh$meta$year <- format(as.Date(out_mbh$meta$date), "%Y")

# extract Topic Proportions (theta) and merge with Year
topic_year_dist <- data.frame(mbh_stm$theta)
colnames(topic_year_dist) <- paste0("Topic", 1:22)
topic_year_dist$year <- out_mbh$meta$year

# calculate mean prevalence per year
year_table <- topic_year_dist %>%
  group_by(year) %>%
  summarise(across(starts_with("Topic"), mean)) %>%
  pivot_longer(cols = starts_with("Topic"), names_to = "Topic", values_to = "Prevalence") %>%
  pivot_wider(names_from = year, values_from = Prevalence)

# view
print(year_table)


```




## MBH Topic Heatmap over Time

```{r}
# A quick visual check of Topic vs. Year
library(ggplot2)

year_long <- topic_year_dist %>%
  pivot_longer(cols = starts_with("Topic"), names_to = "Topic", values_to = "Value") %>%
  group_by(year, Topic) %>%
  summarise(Mean_Prevalence = mean(Value), .groups = 'drop')

ggplot(year_long, aes(x = year, y = Topic, fill = Mean_Prevalence)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red") +
  theme_minimal() +
  labs(title = "Topic Heatmap over Time", subtitle = "Are certain topics only emerging late in the sample?")

```

## Baseline Topic Heatmap over Time
```{r}
# Extract Baseline Time Data
base_year_dist <- data.frame(base_stm$theta)
colnames(base_year_dist) <- paste0("Topic", 1:20)
base_year_dist$year <- format(as.Date(out_base$meta$date), "%Y")

base_year_long <- base_year_dist %>%
  pivot_longer(cols = starts_with("Topic"), names_to = "Topic", values_to = "Value") %>%
  group_by(year, Topic) %>%
  summarise(Mean_Prevalence = mean(Value), .groups = 'drop')

ggplot(base_year_long, aes(x = year, y = Topic, fill = Mean_Prevalence)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") + # Blue for Baseline, Red for Scandal
  scale_y_discrete(labels = base_labels) +
  theme_minimal() +
  labs(title = "Baseline Topic Heatmap", x = "Year", y = "Topic Theme")

```

## Aggregate Topic Evolution over Time
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(Polychrome)

# 15 distinct colors
set.seed(123) #ensures same colours everytime run
dist_palette <- createPalette(22, c("#ff0000", "#00ff00", "#0000ff"))
dist_palette <- as.vector(dist_palette)

# prepare data
year_plot_data <- data.frame(mbh_stm$theta)
colnames(year_plot_data) <- paste0("T", 1:22)
year_plot_data$year <- as.numeric(out_mbh$meta$year)

year_summary <- year_plot_data %>%
  group_by(year) %>%
  summarise(across(starts_with("T"), mean)) %>%
  pivot_longer(cols = starts_with("T"), names_to = "Topic", values_to = "Prevalence")

# plotting
ggplot(year_summary, aes(x = year, y = Prevalence, color = Topic, group = Topic)) +
  geom_line(linewidth = 0.8, alpha = 0.8) + # Thick lines help differentiation
  geom_point(size = 1) +
  theme_minimal() +
  # Apply the Polychrome palette
  scale_color_manual(values = dist_palette, labels = topic_labels) + 
  scale_x_continuous(limits = c(2010, 2026), breaks = seq(2010, 2026, 2)) +
  labs(title = "Aggregate Topic Evolution (2010-2026)",
       subtitle = "Colors generated via Polychrome for maximum visual distance",
       x = "Year", 
       y = "Average Topic Proportion") +
  theme(legend.position = "right",
        legend.text = element_text(size = 8),
        panel.grid.minor = element_blank())
```


## Exploratory vs Managerial Over Time
```{r}
regime_data <- data.frame(mbh_stm$theta)
colnames(regime_data) <- paste0("T", 1:22)

regime_data <- regime_data %>%
  mutate(
    Exploratory = T2 + T3 + T6 + T11 + T12 + T16 + T18 + T19 + T21 + T22,
    Managerial  = T4 + T5 + T8 + T13 + T14 + T17 + T20,
    year = as.numeric(format(as.Date(out_mbh$meta$date), "%Y")),
    date = as.Date(out_mbh$meta$date),
    party = out_mbh$meta$party,
    in_government = out_mbh$meta$in_government
  )

regime_summary <- regime_data %>%
  group_by(party, year, in_government) %>%
  summarise(
    Exploratory = mean(Exploratory),
    Managerial = mean(Managerial),
    .groups = 'drop'
  ) %>%
  pivot_longer(cols = c(Exploratory, Managerial), 
               names_to = "Regime", values_to = "Score")

ggplot(regime_summary, aes(x = year, y = Score, color = Regime, linetype = Regime)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  geom_vline(xintercept = 2021, linetype = "dashed", color = "darkgreen", linewidth = 1) +
  facet_wrap(~party) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(2010, 2025, 2)) +
  scale_color_manual(values = c("Exploratory" = "#E41A1C", "Managerial" = "#377EB8")) +
  labs(
    title = "Exploratory and Managerial Regimes in Parliamentary Debate (2010–2025)",
    subtitle = "Dashed line indicates publication of the Mother and Baby Homes Commission Report (2021)",
    x = "Year",
    y = "Average Regime Score",
    caption = "Exploratory: T2,T3,T6,T11,T12,T16,T18,T19,T21,T22 | Managerial: T4,T5,T8,T13,T14,T17,T20"
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

ggsave("R_Figures/regime_over_time.png", width = 12, height = 8, dpi = 300)
ggsave("R_Figures/regime_over_time.pdf", width = 12, height = 8)
```

```{r}
regime_gov_summary <- regime_data %>%
  group_by(year, in_government) %>%
  summarise(
    Exploratory = mean(Exploratory),
    Managerial = mean(Managerial),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(Exploratory, Managerial),
               names_to = "Regime",
               values_to = "Score") %>%
  mutate(Role = ifelse(in_government == 1, "Government", "Opposition"))

ggplot(regime_gov_summary,
       aes(x = year, y = Score, color = Regime, linetype = Role)) +
  geom_line(linewidth = 1.3) +
  geom_point(size = 2) +
  geom_vline(xintercept = 2021, linetype = "dashed", color = "darkgreen", linewidth = 1) +
  theme_minimal() +
  scale_color_manual(values = c("Exploratory" = "#E41A1C", "Managerial" = "#377EB8")) +
  labs(
    title = "Government and Opposition Language Regimes (2010–2025)",
    x = "Year",
    y = "Average Regime Score",
    caption = "Exploratory: T2,T3,T6,T11,T12,T16,T18,T19,T21,T22 | Managerial: T4,T5,T8,T13,T14,T17,T20"
  ) +
  theme(legend.position = "bottom")

ggsave("R_Figures/government_vs_opposition_regimes.png", width = 10, height = 6, dpi = 300)
ggsave("R_Figures/government_vs_opposition_regimes.pdf", width = 10, height = 6)

```



## save csv for Python
```{r}
regime_gov_monthly <- regime_data %>%
  mutate(month = as.Date(format(date, "%Y-%m-01"))) %>%
  group_by(month, in_government) %>%
  summarise(
    Exploratory = mean(Exploratory),
    Managerial  = mean(Managerial),
    .groups     = "drop"
  ) %>%
  pivot_longer(cols = c(Exploratory, Managerial),
               names_to  = "Regime",
               values_to = "Score") %>%
  mutate(Role = ifelse(in_government == 1, "Government", "Opposition"))

write.csv(regime_gov_monthly, "R_data/regime_gov_monthly.csv", row.names = FALSE)
print("Exported regime_gov_monthly.csv")


```

```{r}
# Export topic-level monthly data
topic_gov_monthly <- regime_data %>%
  mutate(month = as.Date(format(date, "%Y-%m-01"))) %>%
  group_by(month, in_government) %>%
  summarise(across(starts_with("T"), mean), .groups = "drop") %>%
  pivot_longer(cols = starts_with("T"), 
               names_to = "Topic_ID", 
               values_to = "Score") %>%
  mutate(
    Role = ifelse(in_government == 1, "Government", "Opposition"),
    Topic_Label = topic_labels[Topic_ID],
    Regime = case_when(
      Topic_ID %in% c("T4","T5","T8","T13","T14","T17", "T20") ~ "Managerial",
      Topic_ID %in% c("T2","T3","T6","T11","T12","T16","T18", "T19","T21", "T22") ~ "Exploratory",
      TRUE ~ "Other"
    )
  )

write.csv(topic_gov_monthly, "R_data/topic_gov_monthly.csv", row.names = FALSE)
print("Exported topic_gov_monthly.csv")

```

Measures Expected Topic Proportion of Topic 17 (Magdalen) for parties that transitioned between Opposition and Government CHANGEEEE
- Institutional Absorption/Muting effect - Habitual Lag(?)
```{r}
# combine model results with metadata - pull the actual topic percentages (theta) and match them with party and status
df_results <- data.frame(
  party = out_mbh$meta$party,
  in_government = out_mbh$meta$in_government,
  date = out_mbh$meta$date,
  topic17_prop = mbh_stm$theta[, 17]  # T17: Magdalen Laundries & Survivor Testimony
)

chameleon_summary <- df_results %>%
  group_by(party, in_government) %>%
  summarise(
    mean_prop = mean(topic17_prop),
    start_date = min(date),
    end_date = max(date),
    count = n(),
    .groups = 'drop'
  ) %>%
  pivot_wider(
    names_from = in_government,
    values_from = c(mean_prop, start_date, end_date, count)
  )

final_table <- chameleon_summary %>%
  filter(!is.na(mean_prop_0) & !is.na(mean_prop_1)) %>%
  mutate(
    Shift = mean_prop_1 - mean_prop_0,
    Opp_Period = paste0(format(start_date_0, "%Y"), "-", format(end_date_0, "%Y")),
    Gov_Period = paste0(format(start_date_1, "%Y"), "-", format(end_date_1, "%Y"))
  ) %>%
  select(party, Opp_Period, mean_prop_0, Gov_Period, mean_prop_1, Shift)

print(final_table)

```
# -----------
# Intra-Rater Reliability
Krippendrof alpha
```{r}
# Create a blind coding sheet - only FREX words, randomised order
set.seed(42)
random_order <- sample(1:22)

coding_sheet <- data.frame(
  Coding_ID = 1:22,
  Topic_Number = random_order,
  FREX_Words = c(
    "babi, mother, home, westbank, counti, orphanag, children",        # T1
    "redress, exclud, survivor, trauma, six, arbitrari, exclus",       # T2
    "fine, gael, fianna, fáil, polit, taoiseach, parti",               # T3
    "excav, site, coron, burial, director, dna, forens",               # T4
    "childcar, alloc, fund, million, increas, earli, programm",        # T5
    "human, right, court, principl, bill, law, jurisdict",             # T6
    "interim, report, forum, commiss, extens, publish, februari",      # T7
    "payment, card, applic, resid, waiver, elig, quirk",               # T8
    "adopt, birth, adopte, certif, illeg, registr, regist",            # T9
    "archiv, databas, data, seal, gdpr, record, copi",                 # T10
    "bethani, trial, mortal, vaccin, infant, die, unmark",             # T11
    "daughter, love, father, marri, lone, marriag, famili",            # T12
    "shall, minut, busi, stage, exceed, seanad, statement",            # T13
    "transport, byrn, notic, senat, mcgrath, burk, pat",               # T14
    "abort, eighth, repeal, pregnanc, termin, referendum, violenc",    # T15
    "get, point, say, situat, thing, know, tri",                       # T16
    "establish, refer, scope, term, inquiri, examin, judg",            # T17
    "grace, asylum, homeless, mental, complaint, sexual, defenc",      # T18
    "victim, shame, truth, justic, horror, incarcer, apolog",          # T19
    "welcom, thank, import, stori, everyon, colleagu, listen",         # T20
    "lei, sin, mar, narrat, ach, utter, read",                         # T21
    "hospit, matern, chariti, etho, church, religi, ownership"         # T22
  )[random_order]
)

write.csv(coding_sheet, "R_Tables/blind_coding_sheet.csv", row.names = FALSE)
print("Blind coding sheet exported — do not look at original labels!")
```

```{r}
library(irr)

# After completing round 2, enter both sets of labels as numeric categories
# convert labels to numbers - Managerial=1, Exploratory=2, Other=3
round1 <- c(3,2,2,1,1,2,1,1,3,1,2,2,1,1,3,2,1,2,2,1,2,2) # the original regime assignments
round2 <- c(3,2,2,1,1,2,1,1,3,1,2,2,1,1,3,2,1,2,2,1,2,2) # fill in after second coding

ratings <- rbind(round1, round2)
kripp.alpha(ratings, method = "nominal")
```


```{r}
print(colnames(out_mbh$meta))
```


```{r}
# Export speech-level theta for Python within-speaker analysis
theta_df <- as.data.frame(mbh_stm$theta)
colnames(theta_df) <- paste0("T", 1:22)

# Add metadata
theta_df$speaker_norm  <- out_mbh$meta$speaker_norm
theta_df$party         <- out_mbh$meta$party
theta_df$in_government <- out_mbh$meta$in_government
theta_df$date          <- out_mbh$meta$date

# Updated regime scores for K=25
theta_df$Managerial  <- theta_df$T4  + theta_df$T5  + 
                        theta_df$T8  + theta_df$T13 + theta_df$T14 + 
                        theta_df$T17 + theta_df$T20

theta_df$Exploratory <- theta_df$T2  + theta_df$T3  + theta_df$T6  + 
                        theta_df$T11 + theta_df$T12 + theta_df$T16 + 
                        theta_df$T18 + theta_df$T19 + theta_df$T21 + 
                        theta_df$T22


write.csv(theta_df, "R_data/theta_speech_level.csv", row.names = FALSE)
print(paste("Exported", nrow(theta_df), "speeches with theta values"))

# Verify regime scores sum sensibly
print(paste("Mean Managerial:", round(mean(theta_df$Managerial), 3)))
print(paste("Mean Exploratory:", round(mean(theta_df$Exploratory), 3)))
```



## Change-Point Analysis



```{r}
# 1. Load data
df_mbh_quarterly  <- read_csv("data/df_mbh_quarterly.csv")
regime_monthly    <- read_csv("R_data/regime_gov_monthly.csv")

# 2. Aggregate STM regime scores to quarterly
regime_quarterly <- regime_monthly %>%
  mutate(
    month   = as.Date(month),
    quarter = as.Date(as.yearqtr(month, format = "%Y-%m-%d"))
  ) %>%
  group_by(quarter, Regime, Role) %>%
  summarise(
    Score = mean(Score, na.rm = TRUE),
    .groups = "drop"
  )

# 3. Pivot wide
regime_wide <- regime_quarterly %>%
  pivot_wider(
    names_from  = c(Regime, Role),
    values_from = Score
  ) %>%
  rename(
    Managerial_Gov  = Managerial_Government,
    Managerial_Opp  = Managerial_Opposition,
    Exploratory_Gov = Exploratory_Government,
    Exploratory_Opp = Exploratory_Opposition
  ) %>%
  mutate(
    Managerial_gap  = Managerial_Gov  - Managerial_Opp,
    Exploratory_gap = Exploratory_Gov - Exploratory_Opp
  ) %>%
  arrange(quarter) %>%
  drop_na()

# 4. Pivot sentiment and similarity wide
df_mbh_quarterly <- df_mbh_quarterly %>%
  mutate(quarter = as.Date(quarter)) %>%
  pivot_wider(
    names_from  = in_government,
    values_from = c(parlasent_sentiment, survivor_similarity, speech_count)
  ) %>%
  rename(
    sent_gov  = parlasent_sentiment_1,
    sent_opp  = parlasent_sentiment_0,
    sim_gov   = survivor_similarity_1,
    sim_opp   = survivor_similarity_0
  ) %>%
  mutate(
    sent_gap = sent_gov - sent_opp,
    sim_gap  = sim_gov  - sim_opp
  ) %>%
  arrange(quarter) %>%
  drop_na()

# 5. Filter to 2014Q1-2024Q4
regime_wide <- regime_wide %>%
  filter(quarter >= as.Date("2014-01-01") &
         quarter <= as.Date("2024-12-31"))

df_mbh_quarterly <- df_mbh_quarterly %>%
  filter(quarter >= as.Date("2014-01-01") &
         quarter <= as.Date("2024-12-31"))

# 6. Interpolate missing quarters
full_quarters <- seq(as.Date("2014-01-01"), as.Date("2024-10-01"), by = "quarter")

interpolate_df <- function(df, date_col, full_dates) {
  df %>%
    right_join(data.frame(quarter = full_dates), by = date_col) %>%
    arrange(quarter) %>%
    mutate(across(where(is.numeric), ~ zoo::na.approx(., rule = 2)))
}

regime_wide      <- interpolate_df(regime_wide, "quarter", full_quarters)
df_mbh_quarterly <- interpolate_df(df_mbh_quarterly, "quarter", full_quarters)

# 7. Check
print(paste("STM quarters:", nrow(regime_wide)))
print(paste("Sentiment quarters:", nrow(df_mbh_quarterly)))
print(paste("Any NaN in regime:", sum(is.na(regime_wide))))
print(paste("Any NaN in quarterly:", sum(is.na(df_mbh_quarterly))))

```

```{r}
set.seed(42)

# Helper function
run_ecp <- function(series, dates, name, sig_lvl = 0.05) {
  mat <- matrix(series, ncol = 1)
  result <- e.divisive(mat, sig.lvl = sig_lvl, R = 499)
  
  cp_indices <- result$estimates
  cp_indices <- cp_indices[cp_indices != 1 & cp_indices <= length(dates)]
  
  if (length(cp_indices) == 0) {
    breaks <- "none detected"
  } else {
    breaks <- paste(format(dates[cp_indices], "%Y"),
                    paste0("Q", ceiling(as.numeric(format(dates[cp_indices], "%m")) / 3)),
                    sep = " ")
  }
  
  cat(sprintf("%-35s → %s\n", name, paste(breaks, collapse = ", ")))
  return(list(result = result, breaks = breaks, indices = cp_indices))
}

# Run ECP at three significance levels
for (sig in c(0.05, 0.10, 0.20)) {
  cat(sprintf("\n=== ECP CHANGE POINT RESULTS (sig.lvl = %.2f) ===\n\n", sig))
  
  run_ecp(regime_wide$Managerial_Gov,   regime_wide$quarter,       "Managerial (Gov)",           sig)
  run_ecp(regime_wide$Managerial_Opp,   regime_wide$quarter,       "Managerial (Opp)",           sig)
  run_ecp(regime_wide$Exploratory_Gov,  regime_wide$quarter,       "Exploratory (Gov)",          sig)
  run_ecp(regime_wide$Exploratory_Opp,  regime_wide$quarter,       "Exploratory (Opp)",          sig)
  run_ecp(regime_wide$Managerial_gap,   regime_wide$quarter,       "Managerial Gap (Gov-Opp)",   sig)
  run_ecp(regime_wide$Exploratory_gap,  regime_wide$quarter,       "Exploratory Gap (Gov-Opp)",  sig)
  run_ecp(df_mbh_quarterly$sim_gov,     df_mbh_quarterly$quarter,  "Survivor Similarity (Gov)",  sig)
  run_ecp(df_mbh_quarterly$sim_opp,     df_mbh_quarterly$quarter,  "Survivor Similarity (Opp)",  sig)
  run_ecp(df_mbh_quarterly$sim_gap,     df_mbh_quarterly$quarter,  "Survivor Similarity Gap",    sig)
  run_ecp(df_mbh_quarterly$sent_gov,    df_mbh_quarterly$quarter,  "Sentiment (Gov)",            sig)
  run_ecp(df_mbh_quarterly$sent_opp,    df_mbh_quarterly$quarter,  "Sentiment (Opp)",            sig)
  run_ecp(df_mbh_quarterly$sent_gap,    df_mbh_quarterly$quarter,  "Sentiment Gap",              sig)
}
```

#Multivariate Change-Point Analysis

```{r}
set.seed(42)

# 1. Build multivariate matrices

# Matrix 1: STM regime scores (all 6 series)
regime_matrix <- regime_wide %>%
  select(Managerial_Gov, Managerial_Opp, 
         Exploratory_Gov, Exploratory_Opp,
         Managerial_gap, Exploratory_gap) %>%
  as.matrix()

# Matrix 2: Similarity and sentiment (all 6 series)
sim_sent_matrix <- df_mbh_quarterly %>%
  select(sim_gov, sim_opp, sim_gap,
         sent_gov, sent_opp, sent_gap) %>%
  as.matrix()

# Matrix 3: All series combined (12 series)
all_matrix <- cbind(regime_matrix, sim_sent_matrix)

# 2. Run multivariate ECP
for (sig in c(0.05, 0.10, 0.20)) {
  cat(sprintf("\n=== MULTIVARIATE ECP (sig.lvl = %.2f) ===\n\n", sig))
  
  # STM regime scores
  result_regime <- e.divisive(regime_matrix, sig.lvl = sig, R = 499)
  cp_regime <- result_regime$estimates
  cp_regime <- cp_regime[cp_regime != 1 & cp_regime <= nrow(regime_wide)]
  breaks_regime <- if (length(cp_regime) == 0) "none detected" else {
    paste(format(regime_wide$quarter[cp_regime], "%Y"),
          paste0("Q", ceiling(as.numeric(format(regime_wide$quarter[cp_regime], "%m")) / 3)),
          sep = " ", collapse = ", ")
  }
  cat(sprintf("%-40s → %s\n", "STM Regime Scores (6 series)", breaks_regime))
  
  # Similarity and sentiment
  result_sim_sent <- e.divisive(sim_sent_matrix, sig.lvl = sig, R = 499)
  cp_sim_sent <- result_sim_sent$estimates
  cp_sim_sent <- cp_sim_sent[cp_sim_sent != 1 & cp_sim_sent <= nrow(df_mbh_quarterly)]
  breaks_sim_sent <- if (length(cp_sim_sent) == 0) "none detected" else {
    paste(format(df_mbh_quarterly$quarter[cp_sim_sent], "%Y"),
          paste0("Q", ceiling(as.numeric(format(df_mbh_quarterly$quarter[cp_sim_sent], "%m")) / 3)),
          sep = " ", collapse = ", ")
  }
  cat(sprintf("%-40s → %s\n", "Similarity & Sentiment (6 series)", breaks_sim_sent))
  
  # All series combined
  result_all <- e.divisive(all_matrix, sig.lvl = sig, R = 499)
  cp_all <- result_all$estimates
  cp_all <- cp_all[cp_all != 1 & cp_all <= nrow(regime_wide)]
  breaks_all <- if (length(cp_all) == 0) "none detected" else {
    paste(format(regime_wide$quarter[cp_all], "%Y"),
          paste0("Q", ceiling(as.numeric(format(regime_wide$quarter[cp_all], "%m")) / 3)),
          sep = " ", collapse = ", ")
  }
  cat(sprintf("%-40s → %s\n", "All Series Combined (12 series)", breaks_all))
}




```
# Figure for Multi Change-Point
```{r}
# 1. Combine all series into long format
# STM regime scores
regime_long <- regime_wide %>%
  select(quarter, Managerial_Gov, Managerial_Opp,
         Exploratory_Gov, Exploratory_Opp) %>%
  pivot_longer(-quarter, names_to = "series", values_to = "value") %>%
  mutate(
    group = case_when(
      grepl("Gov", series) ~ "Government",
      TRUE ~ "Opposition"
    ),
    measure = case_when(
      grepl("Managerial", series) ~ "Managerial Score",
      TRUE ~ "Exploratory Score"
    )
  )


sim_quarterly <- df_mbh_quarterly %>%
  select(quarter, sim_opp, sim_gov, sim_gap) %>%
  arrange(quarter)

sent_quarterly <- df_mbh_quarterly %>%
  select(quarter, sent_opp, sent_gov, sent_gap) %>%
  arrange(quarter)

sim_quarterly$quarter  <- as.Date(sim_quarterly$quarter)
sent_quarterly$quarter <- as.Date(sent_quarterly$quarter)
regime_wide$quarter    <- as.Date(regime_wide$quarter)

# Similarity scores
sim_long <- sim_quarterly %>%
  select(quarter, sim_gov, sim_opp) %>%
  pivot_longer(-quarter, names_to = "series", values_to = "value") %>%
  mutate(
    group = case_when(series == "sim_gov" ~ "Government", TRUE ~ "Opposition"),
    measure = "Survivor Similarity",
    value = value * 100  # scale to match existing plot
  )

# Sentiment scores
sent_long <- sent_quarterly %>%
  select(quarter, sent_gov, sent_opp) %>%
  pivot_longer(-quarter, names_to = "series", values_to = "value") %>%
  mutate(
    group = case_when(series == "sent_gov" ~ "Government", TRUE ~ "Opposition"),
    measure = "ParlaSent Sentiment"
  )

# Combine
all_long <- bind_rows(regime_long, sim_long, sent_long) %>%
  mutate(
    measure = factor(measure, levels = c(
      "Managerial Score", "Exploratory Score",
      "Survivor Similarity", "ParlaSent Sentiment"
    )),
    group = factor(group, levels = c("Government", "Opposition"))
  )

```


```{r}
# 2. Plot
p <- ggplot(all_long, aes(x = quarter, y = value,
                           colour = group, group = group)) +
  # Raw quarterly points (faint)
  geom_point(alpha = 0.2, size = 0.8) +
  # Raw line (faint)
  geom_line(alpha = 0.2, linewidth = 0.4) +
  # LOWESS smoothed line
  geom_smooth(method = "loess", span = 0.4, se = FALSE,
              linewidth = 1.2) +
  # Commission Report vertical line
  geom_vline(xintercept = as.numeric(as.Date("2021-01-12")),
             linetype = "dashed", colour = "grey40",
             linewidth = 0.7) +
  annotate("text",
           x = as.Date("2021-01-12"),
           y = Inf, label = "Commission Report",
           hjust = -0.05, vjust = 1.5,
           size = 2.8, colour = "grey40") +
  # Facet by measure
  facet_wrap(~ measure, scales = "free_y", ncol = 2) +
  # Colours matching existing plots
  scale_colour_manual(
    values = c("Government" = "#2166AC", "Opposition" = "#D6604D"),
    name = "Role"
  ) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  labs(
    title = "Quarterly Discourse Measures: Government vs Opposition (2014-2024)",
    x = NULL,
    y = "Score",
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold", size = 10),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  )

ggsave("R_figures/quarterly_discourse_trajectories.png",
       p, width = 12, height = 8, dpi = 300)
ggsave("R_figures/quarterly_discourse_trajectories.pdf",
       p, width = 12, height = 8)

print(p)

```



# Multivariate Mixed Effects Model (Phase 3)
```{r}
# 1. Load data
all_combined <- read_csv("data/all_switchers_combined.csv")

print(paste("Speeches:", nrow(all_combined)))
print(paste("Speakers:", n_distinct(all_combined$speaker_norm)))
print(table(all_combined$in_government))

# 2. Standardise outcomes
# Essential so all four outcomes are on comparable scales
all_combined <- all_combined %>%
  mutate(
    Managerial_z           = scale(Managerial),
    Exploratory_z          = scale(Exploratory),
    parlasent_sentiment_z  = scale(parlasent_sentiment_sent),
    survivor_similarity_z  = scale(survivor_similarity_sent),
    in_government          = as.factor(in_government)
  )

# 3. Reshape to long format
all_long <- all_combined %>%
  pivot_longer(
    cols      = c(Managerial_z, Exploratory_z, 
                  parlasent_sentiment_z, survivor_similarity_z),
    names_to  = "outcome",
    values_to = "value"
  ) %>%
  mutate(
    outcome = recode(outcome,
      "Managerial_z"          = "Managerial",
      "Exploratory_z"         = "Exploratory",
      "parlasent_sentiment_z" = "Sentiment",
      "survivor_similarity_z" = "Similarity"
    ),
    outcome = as.factor(outcome),
    outcome = relevel(outcome, ref = "Managerial")  # Managerial as reference
  )

print(paste("Long format rows:", nrow(all_long)))

# 4. Multivariate mixed model
# in_government * outcome tests whether effect of government status
# differs across the four outcomes
# (1 | speaker_norm) = random intercept per speaker
# (1 | speaker_norm:outcome) = random intercept per speaker-outcome combination

m_multi <- lmer(
  value ~ in_government * outcome + 
    (1 | speaker_norm) + 
    (1 | speaker_norm:outcome),
  data = all_long,
  REML = TRUE
)

summary(m_multi)
```

```{r}
# 5. Extract and present results
tidy_results <- tidy(m_multi, effects = "fixed", conf.int = TRUE) %>%
  filter(grepl("in_government", term)) %>%
  mutate(
    sig = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      TRUE            ~ ""
    ),
    across(where(is.numeric), ~ round(.x, 4))
  ) %>%
  select(term, estimate, std.error, conf.low, conf.high, p.value, sig)

print(tidy_results)
```





